---
title: "homework-week-4"
author: "Maria C. Codlin"
date: "September 27, 2017"
output: html_document
---

#Homework 2

Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. 
  Based on previous knowledge, she believes that the mean number calls she will hear in that time is exactly 15. 
  Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

What is the probability that she will hear more than 8 calls during any given session?
What is the probability that she will hear no calls in a session?
What is the probability that she will hear exactly 3 calls in a session?
Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

#Answers
##What is the probability that she will hear more than 8 calls during any given session?
```{r}
round(ppois(8, lambda=15, lower.tail=FALSE), 3)
```
##What is the probability that she will hear no calls in a session?
```{r}
round(dpois(0, lambda=15),7)
```
##What is the probability that she will hear exactly 3 calls in a session?
```{r}
round(dpois(3, lambda=15),5) #dpois is exact while ppois is with tail
```
##Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30
```{r}
barplot(dpois(x=0:30, lambda=15), ylim=c(0,0.15), space=0, names.arg=0:30, ylab="Probability", xlab="# of calls", col=rainbow(50))
```

##Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions). Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

```{r}
sim<-rpois(n=104, lambda=15)
hist(sim, xlim=c(0,30), ylab="Frequency", xlab="# calls in observation session", main="Simulated frequency of call counts", col="aquamarine") 
```
 
This histogram shows the simulated frequency of observed calls over a two year period. It seems to be a fair approximation of the Poisson Mass distribution above and so it it would likely be an appropriate statistic to apply to this type of data. 
